\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2020-10-23 manuscript.Rnw
Switch from R Markdown to Sweave

---
documentclass: jss
classoption: nojss
citation_package: natbib
bibliography: references.bib
author:
  - name: Devin Incerti
    affiliation: Genentech
    # use this syntax to add text on several lines
    address: |
      | South San Francisco
    email: \email{incerti.devin@gene.com}
  - name: Jeroen Jansen
    affiliation: |
      | University of California, San Francisco
    address: |
      | San Francisco
    email: \email{jeroen.jansen@ucsf.edu}
title:
  plain:     "hesim: Health Economic Simulation Modeling and Decision Analysis"
  formatted: "\\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis"
abstract: >
  To do.
keywords:
  formatted: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, "\\proglang{R}"]
  plain: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
knitr::opts_chunk$set(eval = TRUE)
```

# Introduction
Health technology assessment (HTA) is a systematic approach for comparing competing health technologies to inform the efficient use of health care resources. Publicly funded health systems such as those in Australia, Canada, and the United Kingdom, among others, use HTA to help maximize health gains for the population given a fixed budget. HTA is also commonly used by private payers to guide coverage decisions and the adoption of new technologies \citep{trosman2011health}. Such assessments usually rely heavily on cost-effectiveness analysis (CEA) so that decisions can be made on the basis of a formal evaluation of costs and effects \citep{dakin2015influence}. CEA is made feasible by development of health economic models that simulate costs and effects over relevant time horizons using the totality of available evidence.

@@

`hesim` is an \proglang{R} package that advances the functionality and performance of the existing software. Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. To maximize flexibility and faciliate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources. So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized. After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.

>>>>

\documentclass[article, nojss]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Devin Incerti\\Genentech
   \And Jeroen Jansen\\University of California, San Francisco}
\Plainauthor{Devin Incerti, Jeroen Jansen}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis}
\Plaintitle{hesim: Health Economic Simulation Modeling and Decision Analysis}
%\Shorttitle{A Short Demo Article in \proglang{R}}

%% - \Abstract{} almost as usual
\Abstract{
  To do.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, \proglang{R}}
\Plainkeywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Devin Incerti\\
  Genentech\\
  South San Francisco\\
  E-mail: \email{incerti.devin@gene.com}\\

  Jeroen Jansen\\
  University of California, San Francisco\\
  San Francisco\\
  E-mail: \email{jeroen.jansen@ucsf.edu}\\
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section{Introduction} \label{sec:intro}
Health technology assessment (HTA) is a systematic approach for comparing competing health technologies to inform the efficient use of health care resources. Publicly funded health systems such as those in Australia, Canada, and the United Kingdom, among others, use HTA to help maximize health gains for the population given a fixed budget. HTA is also commonly used by private payers to guide coverage decisions and the adoption of new technologies \citep{trosman2011health}. Such assessments usually rely heavily on cost-effectiveness analysis (CEA) so that decisions can be made on the basis of a formal evaluation of costs and effects \citep{dakin2015influence}. CEA is made feasible by development of health economic models that simulate costs and effects over relevant time horizons using the totality of available evidence.

@@

\pkg{hesim} is an \proglang{R} package that advances the functionality and performance of the existing software. Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. To maximize flexibility and faciliate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources. So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized. After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.

\section{Framework} \label{sec:framework}

\begin{figure}[h]
\centering
\includegraphics[width = 15cm,  keepaspectratio]{econ-eval-process-hesim.png}
\caption{Economic modeling process}\label{fig:econ-eval-process-hesim}
\end{figure}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.
\bibliography{references}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

%% -----------------------------------------------------------------------------

\end{document}
